\chapter{Introduction}
\label{chap:devguide-introduction}

This part of the \RA{} documentation is meant for software developers who want to change or
extend business logic, or simply add new features to the \RA{} application. To develop code
which is maintainable and fits well with the \PO{} framework, it helps to understand the
basic concepts which are described in Chapter~\ref{chap:refguide-concepts}.

Adopters of \RA{} are free to modify source code, e.g. to add features for their own or a
client's use. \RA{} is released under the GPL license, and hence, anything which is
compliant with this license is allowed.

If you would like some of your changes to be included in an official release of \RA{},
please get in touch with a core team member.%
% todo: someone (other than bgi) decide if they like this comment:
\footnote{Code contribution to an open source project is by necessity discretionary, in
both directions: contributors should not reveal information that should remain private to
their enterprise, such as information from (or about) their clients, while the core
project team must exercise some degree of control to maintain the goals of the project,
such as commonality or performance requirements.}


\chapter{Development environment}
\label{chap:devEnvironment}

First you need the source code which can be downloaded from the \PO{} website,
\url{www.pillarone.org}.  Apart from the business logic which is completely written by the
core team and the community members, \RA{} is built on top of leading software frameworks.
The most prominent one is Grails%
	\footnote{Grails\index{Grails|textbf}
		is the open-source web application framework based on Groovy\index{Groovy}.
		See their \href{http://grails.codehaus.org/}{homepage} for more information.},
which itself contains the Spring framework, Hibernate and ULC, to mention the most
important ones.  You don't need to download and install these frameworks separately.
The \RA{} source distribution contains the appropriate versions of these frameworks.

To download the source distribution\ldots

Developing with such a sizable software system is more than just editing files. Hence, we
recommend that you set up an integrated development environment (IDE). The core team uses
IntelliJ IDEA and consequently the source distribution contains the project files for
setting up the complete \RA{} source code in IDEA. The project can also be set up in
Netbeans or Eclipse.

For source code versioning we use \href{http://git-scm.com/}{Git}. Anyone can obtain read
access, by following the steps below.
First you should configure your Git username and password for the project:

\begin{itemize}
	\item[]
		\code{git config user.name "Your Name"}\\*
		\code{git config user.email "your.email@email.com"}
\end{itemize}

Git usernames can also be configured globally, but (at least on Windows) this is not
recognized when committing through IDEA.

When the project is started for the first time, IDEA might complain that path variables are not set.
They have to be set once globally and are necessary to allow for stable project files.
If IDEA does not complain, they can be set at \term{File $\rightarrow$ Settings $\rightarrow$ Path variables}.

\begin{itemize}
	\item PILLARONE\_GRAILS: Should be set to the location where the modified Grails 1.2.0
		version was checked out (see above)
	\item USER\_HOME: Should be set to your user home direcotry (where the .grails folder is)
\end{itemize}


\chapter{Modularization}
\label{chap:devguide-modularization}

\RA{} is split up in different modules, currently core, application and business logic.
As \RA{} is a Grails application, a module is actually a Grails plugin.

For our own and modified public plugins we use our own Grails plugin repository, which is
located at: \url{https://svn.intuitive-collaboration.com/GrailsPlugins/}

Read only access does not require authentication, but a SVN account is required to release
plugins into our repository.


\chapter{Working on existing plugins}
\label{chap:devguide-plugins}

Our plugins are based on a modified version of Grails 1.2.0, which is also available from Git.
\code{ gitaccess@svn.intuitive-collaboration.com:riskanalytics-grails.git
}

The project files of the plugins are already configured correctly to use this grails
version, however you will need to set GRAILS\_HOME to the correct folder if you want to use
grails commands on the CLI.

The core and application plugins can be obtained from our Git repository:
\begin{itemize}
	\item[]
		\code{git clone git://github.com/pillarone/risk-analytics-core.git}\\*
		\code{git clone git://github.com/pillarone/risk-analytics-application.git}
\end{itemize}


Additional plugins are no longer checked into the repository, because we have our own plugin
repository now.  Plugins are defined in application.properties and will be downloaded
automatically.

Changes to the plugin can be commited and pushed to \term{git} as usual. However it is only
possible to change files inside the plugin scope. For example, the application plugin cannot
change anything in \\ \texttt{riskanalytics.core.*}.  If such a change is required, a new
version of the core plugin must be released and then installed in the application plugin.

\section{Releasing a plugin}

At first the distribution location (Grails plugin repository to save the plugin) must be defined.
The discovery location, which is used to download plugins, is already defined in \ixconffile{BuildConfig.groovy}.

To avoid to accidently commit SVN credentials the distribution location should not be defined at the same place.
Instead a \texttt{~/.grails/settings.groovy} file should be created and the following line added:

\code{
grails.plugin.repos.distribution.pillarone = \\*
"https://username:password@svn.intuitive-collaboration.com/GrailsPlugins/"
}

Before releasing, the plugin version number has to be corrected in the file \texttt{RiskAnalytics*GrailsPlugin.groovy}.
If everything is set up, the plugin can be released using the \texttt{release-plugin} ant target.\\

It is probably a good idea to tag the Plugin release in git:
\code{
git tag -m `tagged plugin version x.y' `v.n.n'
}

\section{Running it all together}

As the development is now split up between the different plugins, developing the UI means
that there are only simple test models available and there is no UI for developing business
logic.
This should encourage to write test cases instead.
However sometimes it is necessary to run for example a business logic model with the UI
available. For this case, a `development runner' project is available from git.

\code{
git clone gitaccess@svn.intuitive-collaboration.com:riskanalytics-devrunner.git
}\\

The external dependencies required by core and application are already included.
The location of the \RA{} plugins to be used can be configured in
\ixconffile{BuildConfig.groovy}.
That way it is possible to run and test a complete application, which uses the current
plugin `Snapshots' contrary to the plugin projects which use only released plugin versions
and usually only depend on the core plugin.

However this project should only be used as `application runner' and NOT to edit files of
other plugins!  All changes and running tests should be done in the individal plugin project.


\chapter{Creating your own plugin}
\label{chap:devguide-create-plugins}

All grails plugins are based on a Grails 1.3.4. However due to a bug in the Groovy 1.7.4 compiler, we replaced it with
Groovy 1.7.5. You should therefore use
that version which can be obtained from:
\code{
gitaccess@svn.intuitive-collaboration.com:riskanalytics-grails.git
}

Make sure the environment variable \texttt{GRAILS\_HOME} points to the above mentioned
grails and \texttt{PATH} includes \texttt{GRAILS\_HOME/bin}.
Then simply run \code{grails create-plugin PluginName} to create a new plugin.

\subsection{Define \PO{} repository}

To be able to download from our plugin repository, add the following lines to
\ixconffile{BuildConfig.groovy}:
\begin{verbatim}
grails.project.dependency.resolution = {
    inherits "global" // inherit Grails' default dependencies
    log "warn"

    repositories {
        grailsHome()
        grailsCentral()
    }

    def myResolver = new URLResolver()
    myResolver.addArtifactPattern "https://build.intuitive-collaboration.com/plugins/[artifact]/grails-[artifact]-[revision].[ext]"

    resolver myResolver
}
\end{verbatim}

Most plugins will need the risk-analytics-core plugin:
\code{
grails install-plugin risk-analytics-core
}
The core plugin and all its dependencies will then be downloaded and installed.
Also the file YourPluginGrailsPlugin.groovy should be edited, to define name, version, dependencies and more.
An example *GrailsPlugin.groovy and an example build.xml file can be obtained from the Core or Application plugin.

It is also recommended to update the Config.groovy \& DataSource.groovy files, because certain options should be set.
The easiest way to do this is to look at the corresponding files in the core \& application projects and take the necessary options from there.

\section{Git Hints}

\begin{itemize}
	\item Working on a branch
	\begin{itemize}
		\item create the branch locally using \code{git branch --track 0.5.x remotes/origin/0.5.x}
		\item switch the working branch using \code{git checkout 0.5.x}
	\end{itemize}
\end{itemize}

\section{Environments}

As Risk Analytics is a Grails application we also use Grails' support for different environments. The environment Risk Analytics runs in is defined at startup with the parameter \texttt{-Dgrails.env=environment-name}.
We use the environments for the database connection settings (in grails-app/conf/DataSource.groovy) and several other application settings in grails-app/conf/Config.groovy.

It is possible to add new environments by editing these files and copy-pasting existing environments. This is typically done when a new database product will be used.

For more information see: \href{http://grails.org/doc/latest/guide/3. Configuration.html\#3.2 Environments}{http://grails.org/doc/latest/guide/3.~Configuration.html\#3.2~Environments}

\section{User Management}

The server version of RiskAnalytics includes a user management, which includes login, saving of user preferences and the user information is linked to modelling items.
The user management is based on Spring Security and there are two pre-defined roles, ROLE\_USER who can use the application and ROLE\_ADMIN who can additionaly manage the users.

New users can be added in the BootStrap of your application. An example of how to add new users can be found in the class CoreBootStrap in the core plugin.

\chapter{Scalability}

The principle behind stochastic simulations as applied in RiskAnalytics is to generate a large number of so called independent, identically distributed
realizations of the future ('iteration') and then to statistically analyze the ensemble of all these iterations.

The accuracy (confidence) of the results can be improved by increasing the number of iterations, i.e. the
ensemble the statistical estimates are based on. Typically, asking for better accuracy of the results implies
longer runtimes. This is not necessarily true if
\begin{itemize}
 \item a larger number of computing resources (CPU's) is available and
 \item the software used to run the simulation is capable of distributing the computations to many CPU's.
\end{itemize}

Stochastic simulations are an ideal candidate to distribute the computations since independent iterations can
be generated just as well on different CPU's.

\section{Application Structurue Revisited}
\label{sec:devguide-app-struct}
The RiskAnalytics platform is split up into three modules:
\begin{itemize}
 \item core which provides a basic infrastructure such as data access and reporting services and for setting up
    and running models;
 \item application which, basically, contains the graphical user interface;
 \item business logic.
\end{itemize}

The SimulationRunner is part of the core module and is the component from which the stochastic simulations are
configured, started and run. The following basic steps are processed in a simulation run:
\begin{itemize}
 \item Pre-simulation actions
  \begin{itemize}
   \item Instantiating a model
   \item Loading of input data
  \end{itemize}
 \item Simulation
  \begin{itemize}
   \item Computing the random scenarios (n iterations with m periods each)
   \item Writing the raw simulation output to a file
  \end{itemize}
 \item Post-simulation actions \footnote{Up to RiskAnalytics 1.1 the raw simulation output was automatically written to the database
       as part of the postsimulation actions. In view of eliminating the bottleneck of the data persistence layer to obtain a scalable
       distributed solution this approach was changed: Now the raw simulation output is written to local files and only the statistics
       output is written to the database. The raw simulation output can be loaded into the database on demand.}
  \begin{itemize}
   \item Loading data from file
   \item Computing the statistics
   \item Storing the statistics output in the database
  \end{itemize}

\end{itemize}

The computational intensive and time consuming part of a simulation run clearly is the simulation itself. In order to achieve scalability it
is sufficient to distribute the computation of the random scenarios. Here it is important to assure that all iterations remain stochastically
independent. Furthermore it is important to obtain reproducibility independent of the computing grid configuration (e.g. number of nodes).
\pagebreak

\section{GridGain}

Since 1.2 scalability is a standard feature of the platform. The application should scale on a single multi-core laptop
as well as on a grid consisting of multiple computing resources. No additional deployment overhead is needed; once set-up
and configured, the simulations are automatically  distributed.
The framework used for distributing the computations is GridGain\footnote{For more information about GridGain check
\href{http://www.gridgain.com}{www.gridgain.com}}.
GridGain is open source, java-based hence multi-platform and allows simple deployment. It allows to be integrated smoothly
into the existing platform with minimal intrusion. By using a threaded architecture, it scales well on a multi-core
single machine and on multiple computing resources combined to a grid.

\section{Implementation}
In this section a short description of the implementation of the computing and data persistence part in the gridified RiskAnalytics
application is given.

\subsection{Computing}
\label{sec:devguide-computing}
As discussed in \ref{sec:devguide-app-struct} the Simulation Runner, used to start a stochastic simulation, consists of 3 steps. The "Simulation" step
is computing the random scenarios during multiple iterations. These iterations are independet from each other and can therefore
be run in parallel.
In order to achieve this using the GridGain Framework, the following changes and additions to RiskAnalytics have been made:
\begin{itemize}
 \item The total number of iterations is split into blocks. Each block is allocated to a grid job which then processes the iterations.
 \item A grid job is basically an execution unit which can be run on a grid node. Multiple grid jobs run in parallel.
A grid job consists of a simulation configuration, a list of simulation blocks and the execution routine.
The simulation configuration describes all runtime aspects e.g. numberOfIterations, numberOfPeriods, the parameterization, etc.
A simulation block specifies the iterations to be processed and how to handle the random number generation.
The execution routine creates and starts a Simulation Runner.\footnote{As of RA 1.2 the Simulation Runner executes just the pre-simulations
actions and the simulation actions; the post-simulation actions are no longer part of the Simulation Runner and are called later (see \ref{sec:devguide-data}).}
 \item There is actually one grid node embedded in the Riskanalytics application. This grid node is always active and acts as
master node. The master node is responsible for creating jobs and distributing them to other grid nodes (slave nodes).
These slave nodes process the iterations included in the assigned blocks and send the intermediate results back to the master node.
The master node then also manages the data persistence (see \ref{sec:devguide-data}).
 \item The number of created grid jobs per simulation run is chosen equal to the number of cpus (or cores) in a grid.
\end{itemize}

\subsection{Data persistence}
\label{sec:devguide-data}
Following changes and additions have been made to support the local file persistence:
\begin{itemize}
 \item The master node is continuously storing the received intermediate simulation results.
 \item This raw data consisting of iteration meta information and simulated values is no longer automatically stored into a database
but directly written in binary form into files.
 \item As soon as all grid jobs have finished their simulation run, the reduce method of the master node is called. This method
processes the statistical analysis of the raw data (post simulation process) and stores the statistical characteristics into a database.
\end{itemize}

\section{Configuration}
A grid consists of several computing resources on which a distributed task can be executed. To allow this distributed
computing, a considerable amount of communication and coordination between the computing resources is necessary.
GridGain provides the necessary communication and coordination out-of-the-box, meaning no additional configuration effort
is needed to add grid nodes to an existing grid. The communication is done via network broadcasting, so each grid node
in the same subnet automatically joins the grid.


As mentioned in \ref{sec:devguide-computing} there is already a grid node embedded in the RiskAnalytics application.
When executing a simulation run, this master node is executing the application multi-threaded by considering all available cpu cores
of its host system.
If an additional grid node should join the grid make sure the host system is running in the same subnet as
the Riskanalytics application system and follow these instructions to set up a GridGain node:
\begin{enumerate}
 \item Unzip \texttt{gridgain-3.0.0c-win.zip}
 \item Set \texttt{GRIDGAIN\_HOME} environment variable to the path you unzipped (e.g. \texttt{\path{C:\gridgain-3.0.0c-win}})
 \item Copy the Riskanalytics library \texttt{RiskAnalytics.jar} to \\ \texttt{\path{gridgain-3.0.0c-win\libs\ext}}
 \item Adjust \texttt{\path{gridgain-3.0.0c-win\config\default-spring.xml}}\\
Replace:
\begin{lstlisting}
         <!--
        <property name="p2PLocalClassPathExclude">
            <list>
                <value>org.springframework.*</value>
                <value>org.openspaces.*</value>
                <value>org.hibernate.*</value>
            </list>
        </property>
        -->
\end{lstlisting}

with
\begin{lstlisting}
         <property name="p2PLocalClassPathExclude">
            <list>
                <value>org.apache.commons.logging.*</value>
                <value>org.slf4j.*</value>
            </list>
        </property>
\end{lstlisting}

\item Execute \path{gridgain-3.0.0c-win\bin\ggstart.bat}
A successful start shows on the prompt:\\
\texttt{>>> GridGain started OK}
\item Start the Riskanalytics application. On the prompt of the gridgain node following message should be visible (as soon
as the application has started): \\
\texttt{>>> Node JOINED [nodeId8=..., addr=[...], CPUs=...]}\\
\texttt{>>> Topology snapshot [nodes=2, CPUs=...]}

\item When you start a simulation run with at least 2000 iterations (so at least 2 jobs are generated and can be distributed), you should be able to follow the log output of this simulation run
on the GridGain prompt.

\end{enumerate}
